---
title: "MD population definition"
output: html_document
editor_options: 
  chunk_output_type: console
---
This document is a supplementary file for article:

Genetic heterogeneity and subtypes of major depression

Thuy-Dung Nguyen, M.Sc.(1,2); Arvid Harder, B.Sc.(1); Ying Xiong, M.Sc.(1); Kaarina Kowalec, Ph.D.(1,3); Sara Hägg, Ph.D.(1); Ralf Kuja-Halkola, Ph.D.(1); Christina Dalman, M.D.(2); Patrick F Sullivan, M.D. FRANZCP(1,4); Yi Lu, Ph.D.(1,2)

(1)	Department of Medical Epidemiology and Biostatistics, Karolinska Institutet, Stockholm, Sweden
(2)	Department of Global Public Health, Karolinska Institute, Stockholm, Sweden
(3)	College of Pharmacy, University of Manitoba, Winnipeg, Canada
(4)	Department of Genetics and Psychiatry, University of North Carolina, Chapel Hill, NC, USA

Corresponding author: 
Yi Lu, Ph.D.
Nobels Väg 12A, 
Solna, Sweden 17177  
lu.yi@ki.se 
Responsible for this code file
Thuy-Dung Nguyen
Nobels Väg 12A, 
Solna, Sweden 17177  
thuy.dung.nguyen@ki.se

In this document, we presents R code to derive major depression cases, controls, and exclusion criteria using UK Biobank data (UKB)
# Load packages
```{r setup, include=FALSE, echo=FALSE}
library(data.table)
library(tidyverse)
```

# Load data
Raw data was extracted from UKB phenotype data, the numbers appeared in field names are corresponding to the UKB field ID. Details information and meta-data of UKB can be found here: http://biobank.ctsu.ox.ac.uk/crystal/

List of medication code is here
Ref: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6877131/
We extracted a list of medication as in Supplementary table S2.4 with class 1 = antidepressant, 2 = anti-psychotics, 3 = lithium.

The datasets (.tab file) mentioned below were extracted from 2 data files named ukb21472.tab (date 180427) and ukb22140.tab (date 180524) in a server. To know which variables are stored in each .tab file, see the extraction code in UKB_variable_extraction.Rmd.

```{r, echo=FALSE}
path180427 <- "<path>/ukb180427_21472_p0.tab"
raw180427<- fread(path180427, header = TRUE)
name_180427 = names(raw180427)
names(raw180427)[1] = "eid"
raw180427[,2:7][raw180427[,2:7] == -3 |raw180427[,2:7] == -1] <- NA

path20445_20544 = "<path>/ukb180524_20445_20544.tab"
rawdf2<- fread(path20445_20544, header = TRUE)
names(rawdf2)[1] = "eid"
rawdf2[,2:18][rawdf2[,2:18] == -818 |rawdf2[,2:18] == -313|rawdf2[,2:18] == -121|rawdf2[,2:18] == -819] <- NA

pathdrug <- "<path>/UKB_f20003_antidep_psy_lithi.txt"
dfdrug <- fread(pathdrug, header = T)

path20003 <- "<path>/ukb180427_f20003.tab"
raw20003 <- fread(path20003, header = T)
names(raw20003)[1] = "eid"

path20441_20446_20433 <- "<path>/ukb180524_20441_20446_20433.tab"
raw20441_20446_20433 <- fread(path20441_20446_20433, header = T)
names(raw20441_20446_20433)[1] = "eid"
raw20441_20446_20433[,2:4][raw20441_20446_20433[,2:4] == -121 |raw20441_20446_20433[,2:4] == -818] <- NA

pathcidi  <- "<path>/ukb180524_CIDI_criteria.tab"
rawcidi_ab <- fread(pathcidi, header = T)
names(rawcidi_ab)[1] = "eid"
rawcidi_ab[,2:18][rawcidi_ab[,2:18] == -121 |rawcidi_ab[,2:18] == -818] <- NA

path31 <- "<path>/ukb180427_31.csv"
raw31 <- fread(path31, header = T)
colnames(raw31) = c("eid", "sex")

pathbirth <- "<path>/ukb180427_2734.tab"
rawbirth <- fread(pathbirth, header = T)
names(rawbirth)[1] = "eid"
rawbirth[,2:4][rawbirth[,2:4] == -3] <- NA
```

# I. DEFINE MAJOR DEPRESSION CASES
We defined major depression in UKB based on 7 definitions:
1. CIDI-based MDD
2. ICD-coded MDD
3. Probable MD
4. Self-reported MD
5. MD cardinal symptoms
6. Help-seeking MD
7. Antidepressant use
MD15 is the outer merge of 5 definitions 1-5 used in study 0.

***************H***********************H****************
# 1. CIDI-based MDD
Define 2 groups: lifetime and current CIDI MDD. CIDI MDD cases if in at least 1 definition.
This variable is computed based on method used in Na Cai et al. 
https://www.nature.com/articles/s41588-020-0594-5

## Criteria:

Lifetime MDD
- 5/8 criteria A1-A9 with at least one criteria A1 or A2

A1 (UKB 20446) = Yes
A2 (UKB 20441) = Yes

A3 (UKB 20536) = 1/2/3
A4 (UKB 20532) = Yes
A6 (UKB 20449) = Yes
A7 (UKB 20450) = Yes
A8 (UKB 20435) = Yes
A9 (UKB 20437) = Yes

- Plus Impact on normal roles during worst period of depression
B (UKB 20440) =	3 (a lot)

- Plus Not substance abuse
C  (UKB 41202 + 41204) Not codes F10-F19

- Plus Not SCZ or pyschotic	
D (UKB 41202 + 41204)	Not codes F20-F29

- Plus Not Maniac
E (UKB 41202 + 41204) Not codes F30

************
Current MDD: 

- 5/9 criteria with at least one criteria A1 or A2

A1 (UKB 20510) = 4
A2 (UKB 20514) = 4

A3 (UKB 20511) = 4
A4 (UKB 20517) = 4
A5 (UKB 20518) = 4
A6 (UKB 20519) = 4
A7 (UKB 20507) = 4
A8 (UKB 20508) = 4
A9 (UKB 20513) = 4

- Plus Impact on normal roles during worst period of depression
B (UKB 20440) =	3 (a lot)

- Plus Not substance abuse
C  (UKB 41202 + 41204) Not codes F10-F19

- Plus Not SCZ or psychotic	
D (UKB 41202 + 41204)	Not codes F20-F29

- Plus Not Maniac
E (UKB 41202 + 41204) Not codes F30

## Define variables
### Define CDE critera
```{r}
# Subset data to select variable 41202 and 41204
dfcidi_cde <- raw180427 %>%
  select(c(1, 96:910))

# To list all the unique values in the data cols
code_4120x = unique(as.vector(as.matrix(dfcidi_cde[,2:816])))

# Make a list of code in CDE criteria that exist in the data
cde_code = c("F10", "F11", "F12", "F13", "F14", "F15", "F16", "F17", "F18", "F19", "F20", "F21", "F22", "F23", "F24", "F25", "F28", "F29", "F30")

cde_list=c()
for (i in cde_code) {
    cde_list = append(cde_list, (grep(i, code_4120x, fixed=F, value = T)))
}

# Check if there is any cde criteria code in each ID
dfcidi_cde$not_meet_cde <- as.numeric(apply(dfcidi_cde[, 2:816], 1, function(x) any(x %in% cde_list)))

# Check missing
dfcidi_cde$miss <- as.numeric(apply(dfcidi_cde[, 2:816], 1, function(x) all(is.na(x))))

dfcidi_cde$not_meet_cde[dfcidi_cde$not_meet_cde==0 & dfcidi_cde$miss==1] <- NA

# Flip the code so meet_cde = 1 indicate ID that satisfied cde criteria
dfcidi_cde <- dfcidi_cde %>%
mutate(meet_cde = case_when(not_meet_cde== 1 ~ 0,
                                not_meet_cde== 0 ~ 1)) 

# Select variables
dfcidi_cde <- dfcidi_cde %>%
  select(c("eid","meet_cde"))
```

### Define B criteria
- B criteria 20440 = 3

```{r}
# Get variable from rawcidi_ab
dfcidi_b <- rawcidi_ab %>%
  select(c("eid","f.20440.0.0"))

dfcidi_b$meet_b <- ifelse((dfcidi_b$f.20440.0.0 ==3), 1, 0)

dfcidi_b <- dfcidi_b %>%
  select(c("eid","meet_b"))
```
### Prepare df for AB criteria for lifetime and current CIDI

```{r}
#Get variables 20442 and 20446 from df20441_20446_20433, merge with rawcidi_ab
dfcidi_ab <- raw20441_20446_20433 %>%
  select(c("eid", "f.20441.0.0", "f.20446.0.0")) %>%
  merge(., rawcidi_ab, by="eid")

# Rename columns
criteria_list = c("A1_life20446", "A2_life20441", "A3_life20536", "A4_life20532", "A6_life20449", "A7_life20450", "A8_life20435", "A9_life20437", "A1_cur20510", "A2_cur20514", "A3_cur20511", "A4_cur20517", "A5_cur20518", "A6_cur20519", "A7_cur20507", "A8_cur20508", "A9_cur20513")

var_list = c("20446", "20441", "20536", "20532", "20449", "20450", "20435", "20437", "20510", "20514", "20511", "20517", "20518", "20519", "20507", "20508", "20513")

# Print criteria name and var name to check
for (i in 1:length(criteria_list)) {
  print(c(criteria_list[i], var_list[i]))
}

# Replace variable names with corresponding criteria name
name_dfcidi_ab <- names(dfcidi_ab)
for (i in 1:length(criteria_list)) {
  names(dfcidi_ab)[names(dfcidi_ab) == grep(var_list[i], name_dfcidi_ab, fixed=F, value = T)] <- criteria_list[i]
}

dfcidi_a <- dfcidi_ab %>%
  select(c("eid", all_of(criteria_list)))

```

### Define AB criteria for lifetime CIDI

- 5/8 criteria A1-A9 with at least one criteria A1 or A2

A1 (UKB 20446) = Yes
A2 (UKB 20441) = Yes

A3 (UKB 20536) = 1/2/3
A4 (UKB 20532) = Yes
A6 (UKB 20449) = Yes
A7 (UKB 20450) = Yes
A8 (UKB 20435) = Yes
A9 (UKB 20437) = Yes

```{r}
# At least 1 A1 or A2
dfcidi_a$meet_a12_life <- as.numeric(apply(dfcidi_a[, c("A1_life20446","A2_life20441")], 1, function(x) any(x ==1)))

# At least 5 criteria A1-A9
# Recode criteria A3
dfcidi_a$a3_life <- ifelse((dfcidi_ab$A3_life20536==1| dfcidi_a$A3_life20536==2|dfcidi_ab$A3_life20536==3), 1, 0)

# 5/8 criteria A1-A9
dfcidi_a$sum_a1a9_life <- rowSums(dfcidi_a[,c("A1_life20446","A2_life20441","a3_life","A4_life20532","A6_life20449","A7_life20450", "A8_life20435", "A9_life20437")], na.rm = T)

dfcidi_a$meet_a1a9_life <- ifelse((dfcidi_a$sum_a1a9 >= 5), 1, 0)

dfcidi_a$a1a9_life_missing <- as.numeric(apply(dfcidi_a[,c("A1_life20446","A2_life20441","a3_life","A4_life20532","A6_life20449","A7_life20450", "A8_life20435", "A9_life20437")], 1, function(x) all(is.na(x))))

# Recode meet_a1a9_life = NA if all questions A1-A9 are missing
dfcidi_a$meet_a1a9_life[dfcidi_a$a1a9_life_missing==1 & dfcidi_a$meet_a1a9_life==0] <-NA

dfcidi_a$meet_a_life <- ifelse((dfcidi_a$meet_a12_life==1 & dfcidi_a$meet_a1a9_life==1), 1,0)

```

### Define AB criteria for current CIDI
- 5/9 criteria with at least one criteria A1 or A2

A1 (UKB 20510) = 4
A2 (UKB 20514) = 4

A3 (UKB 20511) = 4
A4 (UKB 20517) = 4
A5 (UKB 20518) = 4
A6 (UKB 20519) = 4
A7 (UKB 20507) = 4
A8 (UKB 20508) = 4
A9 (UKB 20513) = 4

```{r}
# Recode variable for a1/a9
dfcidi_a$a1_cur <- ifelse(dfcidi_a$A1_cur20510==4, 1, 0)
dfcidi_a$a2_cur <- ifelse(dfcidi_a$A2_cur20514==4, 1, 0)
dfcidi_a$a3_cur <- ifelse(dfcidi_a$A3_cur20511==4, 1, 0)
dfcidi_a$a4_cur <- ifelse(dfcidi_a$A4_cur20517==4, 1, 0)
dfcidi_a$a5_cur <- ifelse(dfcidi_a$A5_cur20518==4, 1, 0)
dfcidi_a$a6_cur <- ifelse(dfcidi_a$A6_cur20519==4, 1, 0)
dfcidi_a$a7_cur <- ifelse(dfcidi_a$A7_cur20507==4, 1, 0)
dfcidi_a$a8_cur <- ifelse(dfcidi_a$A8_cur20508==4, 1, 0)
dfcidi_a$a9_cur <- ifelse(dfcidi_a$A9_cur20513==4, 1, 0)

# At least 1 A1 or A2 current
dfcidi_a$meet_a12_cur <- as.numeric(apply(dfcidi_a[, c("a1_cur","a2_cur")], 1, function(x) any(x ==1)))

# 5 or more A1-A9
dfcidi_a$sum_a1a9_cur <- rowSums(dfcidi_a[, c("a1_cur","a2_cur","a3_cur","a4_cur",
                                            "a5_cur","a6_cur","a7_cur","a8_cur","a9_cur")], na.rm = T) # Sum score A1-A9

dfcidi_a$meet_a1a9_cur <- ifelse((dfcidi_a$sum_a1a9_cur >= 5), 1, 0)

dfcidi_a$a1a9_cur_missing <- as.numeric(apply(dfcidi_a[, c("a1_cur","a2_cur","a3_cur","a4_cur",
                                            "a5_cur","a6_cur","a7_cur","a8_cur","a9_cur")], 1, function(x) all(is.na(x))))

dfcidi_a$meet_a1a9_cur[dfcidi_a$meet_a1a9_cur==0 & dfcidi_a$a1a9_cur_missing==1] <- NA

dfcidi_a$meet_a_cur <- ifelse((dfcidi_a$meet_a12_cur==1 & dfcidi_a$meet_a1a9_cur==1), 1,0)

# Subset data to get only criteria a
dfcidi_all <- dfcidi_a %>%
  select(c("eid","meet_a_life", "meet_a_cur")) %>%
  merge(., dfcidi_b, by="eid") %>%
  merge(., dfcidi_cde, by="eid")
```

### Define MDD cidi lifetime and current if meet 4 criteria
```{r}
dfcidi_all$cidi_lifetime <- ifelse((dfcidi_all$meet_a_life == 1 & dfcidi_all$meet_b ==1 & dfcidi_all$meet_cde==1 ), 1, 0)

dfcidi_all$cidi_current <- ifelse((dfcidi_all$meet_a_cur == 1 & dfcidi_all$meet_b ==1 & dfcidi_all$meet_cde==1 ), 1, 0)

dfcidi_all$cidi_mdd <- ifelse((dfcidi_all$cidi_lifetime == 1 | dfcidi_all$cidi_current == 1), 1, 0)
```

### Save data
```{r}
dfcidi <- dfcidi_all %>%
  select(c("eid","cidi_lifetime", "cidi_current", "cidi_mdd"))

save(dfcidi,file="<path>/cidi_case.Rda")
```

***************H***********************H********************H************************H*************

# 2. ICD-coded MDD
Primary or secondary diagnosis of a depressive mood disorder from linked hospital admission records
Ref: https://www.nature.com/articles/s41588-020-0594-5/figures/1?proof=true

## Criteria
ICD code F32, F33, F34, F38, F39 for field 41202 (primary diagnosis) or 41204 (secondary diagnosis) 

## Define variable
```{r}
index41202 = grep(41202, name_180427, value = F)
index41204 = grep(41204, name_180427, value = F)

dficd <- select(raw180427,c(1, all_of(index41202), all_of(index41204)))
```

Filter those who have ICD code F32, F33, F34, F38, F39 for field 41202 (primary diagnosis) or 41204 (secondary diagnosis) 
```{r}
# To list all the unique values in the data cols
icd_ukb=unique(as.vector(as.matrix(dficd[,2:816])))

# Just to have a look how the value look like
#grep("F32", icd_ukb, fixed=F, value = T)
# Make a list of code in the data file to include
icd_mdd = c("F32", "F33", "F34", "F38", "F39")
icd_list=c()
for (i in icd_mdd) {
    icd_list = append(icd_list, (grep(i, icd_ukb, fixed=F, value = T)))
}
# Check if there is any MDD code in each ID
dficd$icd_mdd <- as.numeric(apply(dficd[, 2:816], 1, function(x) any(x %in% icd_list)))

# Recode mdd3 into missing if all variables of ICD codes is missing
dficd$miss <- as.numeric(apply(dficd[, 2:816], 1, function(x) all(is.na(x))))

dficd$icd_mdd[dficd$miss==1] <- NA
```

# Save data dficd
```{r}
dficd<- dficd %>%
  select(c(eid, icd_mdd))

save(dficd,file="<path>/icd.Rda")
```

***************H***********************H********************H************************H*************

# 3. Probable MD
Presence and duration of low mood and anhedonia.
Self-reported help-seeking behaviour for mental health difficulties. 
Ref: https://www.nature.com/articles/s41588-020-0594-5/figures/1?proof=true

Variable name in this study: mdd2

## Criteria
  1. “depressed/down for a whole week (UKB: 4598);
Plus 
at least 2 weeks duration (UKB: 4609);
Plus 
ever seen a GP or psychiatrist for nerves, anxiety or depression” (UKB 2090 or 2010)

Or 

  2. “ever anhedonia for a whole week (UKB 4631); 
plus 
at least 2 weeks duration (UKB 5375); 
plus 
ever seen a GP or psychiatrist for nerves, anxiety, or depression” (UKB: 2090 or 2010).

Criteria 1 and 2 is based on derived variable UKB 20126 (code 3/4/5)
ref: http://biobank.ctsu.ox.ac.uk/crystal/crystal/docs/MentalStatesDerivation.pdf

## Define variable probable_md
```{r}
# Subset data
dfprobable <- raw180427 %>%
  select(c("eid", "f.20126.0.0"))

colnames(dfprobable)= c("eid", "v20126")

# compute variable
dfprobable$probable_md <- ifelse((dfprobable$v20126==3 | dfprobable$v20126==4 | dfprobable$v20126==5), 1, 0)
```

## Save data dfprobable
```{r}
dfprobable <- dfprobable %>%
  select(c("eid","probable_md" ))

save(dfprobable,file="<path>/probable.Rda")
```

***************H***********************H********************H************************H*************

# 4. Self-reported MD
Based on Self-reported non-cancer diseases (UKB 20002)
Ref: https://www.nature.com/articles/s41588-02true0-0594-5/figures/1?proof=
and self-reported Mental health problems ever diagnosed by a professional	UKB 20544	Code 11 (Depression)

## Criteria
UKB 20002	= 1286
or 
UKB 20544 = 11 (Depression)

## Define variable

Criteria 1: UKB 20002	= 1286
```{r}
# Get data
index20002 = grep(20002, names(raw180427), value = F)

df20002 <- select(raw180427,c(1,all_of(index20002)))

# To list all the unique values in the data cols
disease=unique(as.vector(as.matrix(df20002[,2:88])))

# Just to have a look how the value look like
#grep("1286", disease, value = T)
# Check if any of the field 20002* has code 1286
df20002$code1286 <- as.numeric(apply(df20002[, 2:88], 1, function(x) any(x==1286)))

df20002$non_miss <- as.numeric(apply(df20002[, 2:88], 1, function(x) any(!is.na(x))))

# Recode code1286 into 0 if the person ever reported any disease (non-miss = 1)
df20002$code1286[df20002$non_miss==1 & is.na(df20002$code1286) ] <- 0
```

Criteria 2: UKB 20544 = 11 (Depression)

```{r}
# Find the column index
index20544 = grep(20544, names(rawdf2), value = F)

# Subset data
df20544 <- select(rawdf2,c(1,all_of(index20544)))

# Screen for code 11 in all variables 20544
df20544$code11 <- as.numeric(apply(df20544[, 2:17], 1, function(x) any(x==11)))

# Recode missing into 0
df20544$non_miss <- as.numeric(apply(df20544[, 2:17], 1, function(x) any(!is.na(x))))

df20544$code11[df20544$non_miss==1 & is.na(df20544$code11)] <- 0
```

Merge 2 criteria
```{r}
dfself <- df20544 %>%
  select(c("eid", "code11")) %>%
  merge (., df20002, by="eid") %>%
  select(c("eid","code1286", "code11"))

dfself$self_rep_md <- ifelse((dfself$code1286==1 | dfself$code11==1), 1, 0)
```

## Save data dfself
```{r}
dfself <- dfself %>%
  select(c("eid", "self_rep_md"))

save(dfself,file="<path>/self_rep.Rda")
```

***************H***********************H********************H************************H*************

# 5. MD cardinal symptoms

Ever had prolonged loss of interest in normal activities (UKB 20441); or prolonged feelings of sadness or depression (UKB 20446)

## Criteria 
Either 20441 = 1 or 20446 = 1

```{r}
dfcardinal <- raw20441_20446_20433 %>%
  select(c(1,3,4)) 
colnames (dfcardinal) <- c("eid", "v20441", "v20446")

dfcardinal$cardinal_md <- ifelse((dfcardinal$v20441==1 | dfcardinal$v20446==1),1,0)

dfcardinal$non_miss <- as.numeric(apply(dfcardinal[, 2:3], 1, function(x) any(!is.na(x))))
dfcardinal$cardinal_md[(is.na(dfcardinal$cardinal_md) & dfcardinal$non_miss==1)] <- 0
```

## Save data dfcardinal
```{r}
dfcardinal <- dfcardinal %>%
  select(c("eid", "cardinal_md"))

save(dfcardinal,file="<path>/cardinal.Rda")
```

***************H***********************H********************H************************H*************

# 6. Help-seeking MD
Self-reported help-seeking behaviour for mental health difficulties. 
Ref: https://www.nature.com/articles/s41588-020-0594-5/figures/1?proof=true

## Criteria mdd1: 
    1. “Yes” to question “Have you ever seen a general practitioner (GP) for nerves, anxiety, tension or depression?" (UKB 2090)
or
    2. “Yes” to question “Have you ever seen a psychiatrist for nerves, anxiety, tension or depression?” (UKB 2010)

## Derive variable
```{r}
# Get data
index2090= grep("2090", name_180427, value = F)
index2100= grep("2100", name_180427, value = F)
dfseek_help <- select(raw180427,c(1, all_of(index2090), all_of(index2100)))

# Index people with code 1 for any field
dfseek_help$seek_help_md <- as.numeric(apply(dfseek_help[, 2:7], 1, function(x) any(x==1)))
# Check if they have answered for any question then NA -> 0
dfseek_help$non_miss <- as.numeric(apply(dfseek_help[, 2:7], 1, function(x) any(!is.na(x))))
# Recode NA into 0 if there is any answer in any fields 2090 2100
dfseek_help$seek_help_md[is.na(dfseek_help$seek_help_md) & dfseek_help$non_miss==1] <- 0
```

## Save data dfseek_help
```{r}
dfseek_help<- dfseek_help %>%
    select(c(eid, seek_help_md))

save(dfseek_help,file="<path>/seek_help.Rda")
```

***************H***********************H********************H************************H*************

# 7. Medication used
Self-reported Treatment/medication code (UKB field 20003)

## Criteria 
antidep: Using antidepressants
antipsy_lithium: Used antipsychotics and/or lithium

# Compute MDD variable medication use
```{r}
# List all code for each drug
antidep = unique(as.vector(as.matrix(dfdrug[(dfdrug$class==1),1])))
antipsy_lithi = unique(as.vector(as.matrix(dfdrug[(dfdrug$class==2|dfdrug$class==3),1])))

# Index people who used each drug
raw20003$antidep <- as.numeric(apply(raw20003[, 2:145], 1, function(x) any(x %in% antidep)))

raw20003$antipsy_lithium <- as.numeric(apply(raw20003[, 2:145], 1, function(x) any(x %in% antipsy_lithi)))

# Real missing = NA
raw20003$miss <- as.numeric(apply(raw20003[, 2:145], 1, function(x) all(is.na(x))))

raw20003$antidep[raw20003$miss==1 & raw20003$antidep==0] <- NA

raw20003$antipsy_lithium[raw20003$miss==1 & raw20003$antipsy_lithium==0] <- NA
```

# Save data medication
```{r}
dfmedication <- raw20003 %>%
  select(c("eid", "antidep", "antipsy_lithium"))

save(dfmedication,file="<path>/medication.Rda")
```

***************H***********************H********************H************************H*************
# Broad MD definition for study 0
md15 includes those with at least one of the 5 definitions 1-5

```{r}
df_md <- dfcidi %>%
  merge (., dficd, by="eid") %>%
  merge (., dfprobable, by="eid") %>%
  merge (., dfself, by="eid") %>%
  merge (., dfcardinal, by ="eid") %>%
  merge (., dfseek_help, by="eid") %>%
  merge (., dfmedication, by="eid")

df_md$md15 <- as.numeric(apply(df_md[,c("cidi_mdd","icd_mdd","probable_md","self_rep_md","cardinal_md"  )], 1, function(x) any(x ==1)))
# Check if they have answered for any definition then NA -> 0
df_md$non_miss15 <- as.numeric(apply(df_md[,c("cidi_mdd","icd_mdd","probable_md","self_rep_md","cardinal_md")], 1, function(x) any(!is.na(x))))

df_md$md15[is.na(df_md$md15) & df_md$non_miss15==1] <- 0
```
## Save data df_md
```{r}
df_md <- df_md %>%
  select(-non_miss15)

save(df_md,file="Z:/<path>/df_md.Rda")
```

***************H***********************H********************H************************H*************
***************H***********************H********************H************************H*************
***************H***********************H********************H************************H*************
This document contains code to extract MDD control population for UKB cohort 
Metadata here
http://biobank.ctsu.ox.ac.uk/crystal/

# Description
There are 3 controls group for MD cases, PPD cases and comorbid anxiety cases

1. Control for MD cases
- Not included in any of 7 MD definitions (CIDI-based MDD, ICD-coded MDD, Probable MD, Self-reported MD, MD cardinal symptoms, Help-seeking MD, Antidepressant use)

2. Controls for PPD
- Not included in any of 7 MD definitions
- Women (UKB 31)
- Have given >= 1 live birth

3. Control for comorbid anxiety MD
- Not included in any of 7 MD definitions
- No anxiety diagnosis F40, F41 (UKB 41202, 41204)

```{r}
load("<path>/df_md.Rda") #df that was created in the previous section
```

# Preprocess data
1. Not cases

```{r}
df_md$control_md <- ifelse((df_md$cidi_mdd==1 | df_md$icd_mdd==1 | df_md$probable_md==1 | df_md$self_rep_md==1 | df_md$cardinal_md== 1 | df_md$seek_help_md==1 | df_md$antidep==1), 0, 1)

df_md$control_md[is.na(df_md$control_md)] <- 1
df_md$control_md[df_md$control_md==0] <- NA
```

2. Indexing mothers
```{r}
# Indexing those without have given birth
rawbirth$birth <- as.numeric(apply(rawbirth[, 2:4], 1, function(x) any(x>0)))
rawbirth$birth[rawbirth$birth ==0] <- NA

dfmother <- rawbirth %>%
  select(c("eid","birth")) %>%
  merge (., raw31, by="eid")

dfmother$mother <- ifelse((dfmother$birth==1 & dfmother$sex==0), 1, 0)
dfmother$mother[dfmother$mother==0] <-NA

dfmother <- dfmother %>%
  select(c("eid","mother"))

table(dfmother$birth, dfmother$sex, useNA = "ifany")
table(dfmother$mother, useNA = "ifany")

```
3. Anxiety

```{r}
# Index those with anxiety diagnosis
index41202 = grep(41202, name_180427, value = F)
index41204 = grep(41204, name_180427, value = F)

dfanxiety <- select(raw180427,c(1, all_of(index41202), all_of(index41204)))

code_4120x = unique(as.vector(as.matrix(dfanxiety[,2:816])))

anx_code = c("F40", "F41")

anx_list=c()
for (i in anx_code) {
    anx_list = append(anx_list, (grep(i, code_4120x, fixed=F, value = T)))
}
# Check if there is any anxiety code in each ID
dfanxiety$anxiety <- as.numeric(apply(dfanxiety[, 2:816], 1, function(x) any(x %in% anx_list)))

# Only exclude those with anxiety diagnosis from control for comorbid anxiety cases, missing are kept

dfanxiety <- dfanxiety %>%
  select(c(eid, anxiety))
```

# Merge data
```{r}
# Subset dataset for criteria 1
dfcontrol <- df_md %>%
  select(c("eid","control_md")) %>%
  merge(., dfmother, by="eid") %>%
  #merge(., dfanxiety, by="eid") %>%
  merge(., raw31, by="eid")

dfcontrol$control_ppd <- ifelse((dfcontrol$control_md==1 & dfcontrol$mother==1), 1, 0)
dfcontrol$control_ppd[dfcontrol$control_ppd==0] <- NA

# Anxiety = 0 are those without anxiety
dfcontrol$control_anxiety <- ifelse((dfcontrol$control_md==1 & dfcontrol$anxiety!=1), 1, 0)
dfcontrol$control_anxiety[dfcontrol$control_anxiety==0] <- NA
```

# Filter data criteria 2 and save data
```{r}
dfcontrol <- dfcontrol %>%
  select(c("eid", "control_md","control_ppd", "control_anxiety"))

save(dfcontrol,file="<path>/dfcontrol.Rda")
```

