---
title: "Subtype derivation"
author: "Thi Thuy Dung Nguyen"
date: '2020-11-01'
output: html_document
editor_options: 
  chunk_output_type: console
---
This document is a supplementary file for the manuscript:

Thuy-Dung Nguyen, M.Sc.(1,2); Arvid Harder, B.Sc.1; Ying Xiong, M.Sc.(1); Kaarina Kowalec, Ph.D.(1,3); Sara H채gg, Ph.D.(1); Na Cai, Ph.D.(4), Ralf Kuja-Halkola, Ph.D.(1); Christina Dalman, M.D.(2); Patrick F Sullivan, M.D. FRANZCP(1,5); Yi Lu, Ph.D.(1,2)

1.	Department of Medical Epidemiology and Biostatistics, Karolinska Institutet, Stockholm, Sweden
2.	Department of Global Public Health, Karolinska Institute, Stockholm, Sweden
3.	College of Pharmacy, University of Manitoba, Winnipeg, Canada
4.	Helmholtz Pioneer Campus, Helmholtz Zentrum M체nchen, Neuherberg, Germany
5.	Department of Genetics and Psychiatry, University of North Carolina, Chapel Hill, NC, USA

Corresponding author: 
Yi Lu, Ph.D.
Nobels V채g 12A, 
Solna, Sweden 17177  
lu.yi@ki.se 
Responsible for this code file
Thuy-Dung Nguyen
Nobels V채g 12A, 
Solna, Sweden 17177  
thuy.dung.nguyen@ki.se

This file contains code to derive MD subtypee. Raw data was extracted from UKB phenotype data, the numbers appeared in field names are corresponding to the UKB field ID. Details information and meta-data of UKB can be found here: http://biobank.ctsu.ox.ac.uk/crystal/

The small datasets (.tab file) mentioned below were extracted from 2 original data files named ukb21472.tab (date 180427) and ukb22140.tab (date 180524). To know which variables are stored in each .tab file, see the extraction code in UKB_variable_extraction.Rmd.

# Loading packages
```{r, include=FALSE}
library(data.table)
library(tidyverse)
```
# Loading data
```{r}
path180427 <- "<path>/ukb180427_21472_p0.tab"
raw180427<- fread(path180427, header = TRUE)
name_180427 = names(raw180427)
names(raw180427)[1] = "eid"
raw180427[,2:7][raw180427[,2:7] == -3 |raw180427[,2:7] == -1] <- NA # 2090 & 2100

path20433 <- "<path>/ukb180524_20441_20446_20433.tab"
raw20433 <- fread(path20433, header = TRUE)
names(raw20433)[1] = "eid"
raw20433[,2:4][raw20433[,2:4] == -121 |raw20433[,2:4] == -818] <- NA

path20534 <- "<path>/ukb180524_20534.tab"
raw20534 <- fread(path20534, header=T)
names(raw20534)[1] = "eid"

pathcidi  <- "<path>/ukb180524_CIDI_criteria.tab"
rawcidi_ab <- fread(pathcidi, header = T)
names(rawcidi_ab)[1] = "eid"
rawcidi_ab[,2:18][rawcidi_ab[,2:18] == -121 |rawcidi_ab[,2:18] == -818] <- NA

path20126 <- "<path>/ukb180427_20126.tab"
raw20126 <- fread(path20126, header = T)
names(raw20126) = c("eid", "v20126")

path20445_20544 = "<path>/ukb180524_20445_20544.tab"
raw20544<- fread(path20445_20544, header = TRUE, drop="f.20445.0.0")
names(raw20544)[1] = "eid"
raw20544[,2:17][raw20544[,2:17] == -819 |raw20544[,2:17] == -818] <- NA

raw20445<- fread(path20445_20544, header = TRUE) %>%
  select(c(f.eid, f.20445.0.0))
names(raw20445)[1] = "eid"
raw20445[,2][raw20445[,2] == -818 |raw20445[,2] == -313|raw20445[,2] == -121] <- NA

path20437_20447 <- "<path>/ukb180524_20437_20447.tab"
raw20437_20447 <- fread(path20437_20447, header = TRUE)
names(raw20437_20447) = c("eid", "v20437", "v20447")
raw20437_20447[,2:3][raw20437_20447[,2:3] == -121 |raw20437_20447[,2:3] == -818] <- NA

path31 <- "<path>/ukb180427_31.csv"
rawsex <- fread(path31, header = T)
colnames(rawsex) = c("eid", "sex")

# Case-Controls: final_pop
load("<path>/case_control_201029.Rda")
```
# Derive subtypes

## 1. Atypical symptoms
```{r}
dfatypical <- rawcidi_ab %>% 
  select(c("eid", "f.20536.0.0")) %>%
  merge (., raw20534, by="eid") %>%
  merge (., final_pop, by="eid") %>%
  select(c("fid", "eid", "cidi_mdd", "md15", "f.20536.0.0", "f.20534.0.0", "control_md"))
```

Computing var symptoms = 1 if (20536 = 1 & 20534 = 1)
= NA if (20536 = NA | 20534 = NA)
= 0 otherwise

```{r}
dfatypical$symptom <- ifelse((dfatypical$f.20536.0.0==1 & dfatypical$f.20534.0.0 ==1), 1,0)
dfatypical$symptom[is.na(dfatypical$symptom)] <- 0
dfatypical$symptom[dfatypical$symptom==0 & (is.na(dfatypical$f.20536.0.0)|is.na(dfatypical$f.20534.0.0))] <- NA
```

### a. Atypical_cardinal
atypical = 2 if (md15 =1 & symptom = 1 (weight gain and Hypersomnia))
= 1 if (md15 =0 & symptom = 0 (weight gain and Hypersomnia))
= NA otherwise (any of those 3 criteria = NA)
= 0 if control

```{r}
dfatypical$atypical <- ifelse(dfatypical$md15==1 & dfatypical$symptom==1, 2, NA)
dfatypical$atypical[is.na(dfatypical$atypical) & dfatypical$md15==1 & dfatypical$symptom==0] <-1

# Control =0
dfatypical$atypical[is.na(dfatypical$atypical) & dfatypical$control_md==1] <- 0

```

### b. Atypical CIDI
Same as atypical but replace md15 = cidi_mdd

```{r}
dfatypical$atypical_cidi <- ifelse(dfatypical$cidi_mdd==1 & dfatypical$symptom==1, 2, NA)

dfatypical$atypical_cidi[is.na(dfatypical$atypical_cidi) & dfatypical$cidi_mdd==1 & dfatypical$symptom==0] <-1

# Control =0
dfatypical$atypical_cidi[is.na(dfatypical$atypical_cidi) & dfatypical$control_md==1] <- 0
```

### Select variables

```{r}
dfatypical <- dfatypical %>%
  select(c("fid","eid", "atypical", "atypical_cidi"))
```

***************HHHHHH****************

## 2. Severity
Severity among cases defined in probable MD and ICD codes

Severity = 1, mild/moderate:

20126=4
or
41202 = F320/F321(single mild/moderate) / F330/F331 (recurrent mild/moderate)
or
41204 = F320/F321(single mild/moderate) / F330/F331 (recurrent mild/moderate)

Severity = 2, severe:

20126=3
or
41202 = F322, F323 (single severe)/ F332, F333 (recurrent severe)
or
41204 = F322, F323 (single severe)/ F332, F333 (recurrent severe)


### Compute severity based on 20126
severe20126 = 1 if v20126 3
severe20126 = 0 if v20126 4

```{r}
dfsevere20126 <- raw20126

dfsevere20126$v20126[dfsevere20126$v20126==5 | dfsevere20126$v20126 <3 ] <- NA

dfsevere20126$v20126[dfsevere20126$v20126==4 ] <- 0 # Moderate
dfsevere20126$v20126[dfsevere20126$v20126==3 ] <- 1 # Severe

names(dfsevere20126) = c("eid","severe20126")
```
### Compute severity based on ICD codes
icdsevere
= 1 if any ICD code in severe_list = c("F322","F323", "F332", "F333")
= 0 if any ICD no code in severe_list but not all missing
= NA if icdmiss = 1 (missing all ICD fields)

```{r}
index41202 = grep(41202, names(raw180427), value = F)
index41204 = grep(41204, names(raw180427), value = F)

df4120x <- select(raw180427,c(1, all_of(index41202), all_of(index41204)))
```

Filter those who have ICD code for field 41202 (primary diagnosis) or 41204 (secondary diagnosis) 

```{r}
# To list all the unique values in the data cols
icd_ukb=unique(as.vector(as.matrix(df4120x[,2:816])))

# To see the list code in UKB
#f33_list <- grep("F33", icd_ukb, fixed=F, value = T)
#mild_list = c("F320","F321","F330", "F331")
severe_list = c("F322","F323", "F332", "F333")

df4120x$icdmiss <- as.numeric(apply(df4120x[, 2:816], 1, function(x) all(is.na(x))))

df4120x$icdsevere <- as.numeric(apply(df4120x[, 2:816], 1, function(x) any(x %in% severe_list)))
# The rest of ICD cases are considered mild moderate
df4120x$icdsevere[df4120x$icdsevere==0 & df4120x$icdmiss ==1] <- NA
```

### Combine the 2 severity conditions
```{r}
dfseverity <- df4120x %>%
  merge (., dfsevere20126, by="eid") %>%
  merge (., final_pop, by="eid") %>%
  select (c("fid", "eid","severe20126", "icdsevere", "md15", "cidi_mdd", "control_md"))
```

severe_ind = 1 if (severe20126 = 1 | icdesevere = 1)
= 0 if (severe20126 = 0 & icdesevere = 0)
= NA if (severe20126 = NA | icdesevere = NA)

```{r}
dfseverity$severe_ind <- ifelse((dfseverity$severe20126==1 | dfseverity$icdsevere ==1), 1, NA)
  
dfseverity$severe_ind[is.na(dfseverity$severe_ind) & dfseverity$severe20126==0 &dfseverity$icdsevere ==0] <- 0
```

severity = 2 if (severe_ind =1 & md15 = 1)
= 1 if (severe_ind =0 & md15 = 1)
= 0 if control
= NA if (severe_ind =NA | md15 = NA)
```{r}
dfseverity$severity <- ifelse((dfseverity$md15==1 & dfseverity$severe_ind ==1), 2, NA)

dfseverity$severity[is.na(dfseverity$severity) & dfseverity$md15==1 & dfseverity$severe_ind ==0] <- 1

dfseverity$severity[is.na(dfseverity$severity) & dfseverity$control_md==1] <- 0
```
severity_cidi = 2 if (severe_ind =1 & cidi_mdd = 1)
= 1 if (severe_ind =0 & cidi_mdd = 1)
= 0 if control
= NA if (severe_ind =NA | md15 = NA)

```{r}
dfseverity$severity_cidi <- ifelse((dfseverity$cidi_mdd==1 & dfseverity$severe_ind ==1), 2, NA)

dfseverity$severity_cidi[is.na(dfseverity$severity_cidi) & dfseverity$cidi_mdd==1 & dfseverity$severe_ind ==0] <- 1

dfseverity$severity_cidi[is.na(dfseverity$severity_cidi) & dfseverity$control_md==1] <- 0
```
### Select variables
```{r}
dfseverity <- dfseverity %>%
  select(c("fid","eid","severity", "severity_cidi"))
```

***************HHHHHH****************
## 3. Comorbid anxiety
MD with comobid anxiety, panic attacks, social anxiety/phobia	
20544 = 1/6/15
or 
41202/41204 = F40, F41 (Phobic anxiety disorder/other anxiety disorder))
Non-comorbid anxiety	
20544 != 1/6/15
and 41202/41204 != F40, F41)

### Filter those with anxiety
Filter those with anxiety in 20544

```{r}
all20544_list =unique(as.vector(as.matrix(raw20544[,2:17])))
axn_list <- c(1,6,15)

raw20544$anx20544 <- as.numeric(apply(raw20544[, 2:17], 1, function(x) any(x %in% axn_list)))

raw20544$miss <- as.numeric(apply(raw20544[, 2:17], 1, function(x) all(is.na(x))))

raw20544$anx20544[raw20544$miss==1& raw20544$anx20544==0] <- NA

```
Anxiety in icd
```{r}
index41202 = grep(41202, names(raw180427), value = F)
index41204 = grep(41204, names(raw180427), value = F)

df4120x <- select(raw180427,c(1, all_of(index41202), all_of(index41204)))
```


Filter those who have ICD code F40, F41 for field 41202 (primary diagnosis) or 41204 (secondary diagnosis) 
```{r}
# To list all the unique values in the data cols
#icd_ukb=unique(as.vector(as.matrix(df4120x[,2:816])))

anxiety_list = c("F401","F400","F402","F408","F409","F410","F419","F412","F411","F418","F413")

df4120x$icdanxiety <- as.numeric(apply(df4120x[, 2:816], 1, function(x) any(x %in% anxiety_list)))

# Check missing in all ICD fields and code = NA, use later when applying cases condition
df4120x$miss <- as.numeric(apply(df4120x[, 2:816], 1, function(x) all(is.na(x))))

df4120x$icdanxiety[df4120x$miss==1] <- NA

```

Computing co_anxiety
= 1 if anx20544 =1 | icdanxiety = 1
= 0 if anx20544 =0 and icdanxiety = 0

```{r}
dfanxiety <- df4120x %>%
  select(c("eid", "icdanxiety")) %>%
  merge(., raw20544, by="eid") %>%
  merge(., final_pop, by="eid") %>%
  select(c("fid", "eid","anx20544","icdanxiety", "md15", "cidi_mdd","control_anxiety"))

dfanxiety$co_anxiety <- ifelse((dfanxiety$anx20544==1 | dfanxiety$icdanxiety==1), 1, NA)

dfanxiety$co_anxiety[is.na(dfanxiety$co_anxiety)& (dfanxiety$anx20544==0 & dfanxiety$icdanxiety==0)] <- 0
```

Compute co.anxiety MD
```{r}
# compute anxiety =2 if md15 == 1 and co_anxiety ==1
dfanxiety$anxiety <- ifelse((dfanxiety$md15==1 & dfanxiety$co_anxiety==1), 2, NA)

dfanxiety$anxiety[is.na(dfanxiety$anxiety) & dfanxiety$md15==1 & dfanxiety$co_anxiety==0] <- 1

dfanxiety$anxiety[is.na(dfanxiety$anxiety) & dfanxiety$control_anxiety ==1] <- 0

# compute anxiety_cidi = 2 if cidi_mdd = 1 and co_anxiety = 1

dfanxiety$anxiety_cidi <- ifelse((dfanxiety$cidi_mdd==1 & dfanxiety$co_anxiety==1), 2, NA)

dfanxiety$anxiety_cidi[is.na(dfanxiety$anxiety_cidi) & dfanxiety$cidi_mdd==1 & dfanxiety$co_anxiety==0] <- 1

dfanxiety$anxiety_cidi[is.na(dfanxiety$anxiety_cidi) & dfanxiety$control_anxiety ==1] <- 0
```

### Select variables

```{r}
dfanxiety <- dfanxiety %>%
  select(c("fid", "eid", "anxiety", "anxiety_cidi"))
```

***************HHHHHH****************
## 4. Age at onset
```{r}
dfaoo <- raw20433 %>%
  merge(., final_pop, by="eid") %>%
  select(c("fid", "eid", "f.20433.0.0", "control_md", "md15", "cidi_mdd")) 
names(dfaoo)[3] <- "v20433"

octile = quantile(dfaoo$v20433, probs = seq(0,1,0.125), na.rm = T, names = F)

dfaoo$aoo= as.character(cut(dfaoo$v20433,octile,right=TRUE, include.lowest=TRUE))

dfaoo$aoo[(dfaoo$aoo=="[2,19]"| dfaoo$aoo== "(19,25]"|dfaoo$aoo== "(25,30]")] <- 1
dfaoo$aoo[(dfaoo$aoo=="(30,37]"| dfaoo$aoo== "(37,43]")] <- NA
dfaoo$aoo[(dfaoo$aoo=="(43,49]"| dfaoo$aoo== "(49,55]"|dfaoo$aoo== "(55,78]")] <- 2

dfaoo$aoo[is.na(dfaoo$aoo) & dfaoo$md15==0] <- 0

dfaoo$aoo_cidi <- dfaoo$aoo
dfaoo$aoo_cidi[(dfaoo$aoo_cidi==1 | dfaoo$aoo_cidi==2) & is.na(dfaoo$cidi_mdd)] <- NA
```

### Select variables
```{r}
dfaoo <- dfaoo %>%
  select(c("fid", "eid","aoo", "aoo_cidi")) 
```

***************HHHHHH****************

## 5. Recurrence
Single episode MDD or CIDI-based single episode	

Single:
20126   = 5 
or 
mdd8_cidi = 1 & 20442 < 2
or 
ICD = F32 (no F33)

Recurrent 
20126   = 3/4 
or 
mdd8_cidi = 1 & 20442 >= 2
or 
ICD = F33

### Compute recur_ind
recur_ind
= 1 if 
20126   = 3/4 
or 
mdd8_cidi = 1 & 20442 >= 2
or 
ICD = F33

=0 if 
20126   = 5 
or 
mdd8_cidi = 1 & 20442 < 2
or 
ICD = F32 (no F33)

```{r}
# Filter ICD codes
index41202 = grep(41202, names(raw180427), value = F)
index41204 = grep(41204, names(raw180427), value = F)

df4120x <- select(raw180427,c(1, all_of(index41202), all_of(index41204)))
# To list all the unique values in the data cols
icd_ukb=unique(as.vector(as.matrix(df4120x[,2:816])))

# List of F33 codes in ukb
#f32_list <- grep("F32", icd_ukb, fixed=F, value = T)
sing_list = c("F329", "F323", "F322", "F321", "F320", "F328", "F32")
recur_list = c("F331", "F333", "F332", "F339", "F334", "F330", "F338")

# icd_recur = 1 if in recur_list
df4120x$icd_recur <- as.numeric(apply(df4120x[, 2:816], 1, function(x) any(x %in% recur_list)))

df4120x$icd_single <- as.numeric(apply(df4120x[, 2:816], 1, function(x) any(x %in% sing_list)))

# if in recur list then single is 0
df4120x$icd_single[df4120x$icd_single==1 & df4120x$icd_recur==1] <- 0

# Function any() indicate 1 if there condition = T and 0 if not. So recode recur at 0 to NA if single = 0
df4120x$icd_recur[df4120x$icd_single==0 & df4120x$icd_recur==0] <- NA

# Combine with other dataset
dfrecurrence <- df4120x %>%
  select(c("eid", "icd_recur")) %>%
  merge(., rawcidi_ab, by="eid") %>%
  merge(., raw20126, by="eid") %>%
  merge(., final_pop, by="eid") %>%
  select(c("fid", "eid", "icd_recur", "v20126", "f.20442.0.0", "cidi_mdd", "md15", "control_md"))
```

Compute cidi recurrent
```{R}
# Recode -999 to 1000 as the case has too many episode to count
dfrecurrence$f.20442.0.0[dfrecurrence$f.20442.0.0 == -999] <- 1000

dfrecurrence$multi_eps <- ifelse((dfrecurrence$f.20442.0.0 >= 2), 1, 0)

# Compute recurrent cidi-based MDD 
dfrecurrence$cidi_recur <- ifelse((dfrecurrence$cidi_mdd == 1 & dfrecurrence$multi_eps ==1), 1, NA)
dfrecurrence$cidi_recur[is.na(dfrecurrence$cidi_recur) & dfrecurrence$cidi_mdd == 1 & dfrecurrence$multi_eps ==0] <- 0
```

Compute 20126 recurrent
```{R}
# Recode recur20126 = 1 if v20125 = 3/4
# recur20126 = 0 if v20125 = 5
# Otherwise NA
dfrecurrence$recur20126 <- NA
dfrecurrence$recur20126[dfrecurrence$v20126 ==3 | dfrecurrence$v20126 == 4] <- 1
dfrecurrence$recur20126[dfrecurrence$v20126 ==5] <- 0
```

Compute recur_ind
=1 if icd_recur =1 | cidi_recur=1 | recur20126=1
=0 if icd_recur =0 | cidi_recur=0 | recur20126=0 but not = 1

```{r}
dfrecurrence$recur_ind <- ifelse((dfrecurrence$icd_recur==1 |dfrecurrence$cidi_recur==1 | dfrecurrence$recur20126==1), 1, NA)

dfrecurrence$recur_ind[is.na(dfrecurrence$recur_ind) & (dfrecurrence$icd_recur==0 |dfrecurrence$cidi_recur==0 | dfrecurrence$recur20126==0)] <- 0
```

Compute recurrence
=1 if recur_ind=1 & md15= 1
=0 if recur_ind=0 & md15= 1

```{r}
# md15
dfrecurrence$recurrence <- ifelse((dfrecurrence$md15==1 & dfrecurrence$recur_ind==1), 2, NA)
dfrecurrence$recurrence[is.na(dfrecurrence$recurrence) & dfrecurrence$md15==1 & dfrecurrence$recur_ind==0] <- 1
dfrecurrence$recurrence[is.na(dfrecurrence$recurrence) & dfrecurrence$control_md==1] <- 0

# Recurrence cidi
dfrecurrence$recurrence_cidi <- ifelse((dfrecurrence$cidi_mdd==1 & dfrecurrence$recur_ind==1), 2, NA)
dfrecurrence$recurrence_cidi[is.na(dfrecurrence$recurrence_cidi) & dfrecurrence$cidi_mdd==1 & dfrecurrence$recur_ind==0] <- 1
dfrecurrence$recurrence_cidi[is.na(dfrecurrence$recurrence_cidi) & dfrecurrence$control_md==1] <- 0
```

### Select variables
```{r}
dfrecurrence <- dfrecurrence %>%
  select(c("fid", "eid","recurrence", "recurrence_cidi"))
```

***************HHHHHH****************
##. 6. Suicide
Having suicidal thought during the worst period of MDD indication or recent thoughts of suicide or self-harm.

Yes (2):
20437 = 1
or
20513 = 3/4

No:
20437 = 0
or
20513 = 1

```{r}
dfsuicide <- raw20437_20447 %>%
  merge (., rawcidi_ab, by= "eid") %>%
  select (c("eid", "v20437","f.20513.0.0")) %>%
  merge(., final_pop, by="eid")

# Recode 20513
dfsuicide$f.20513.0.0[dfsuicide$f.20513.0.0==1] <- 0
dfsuicide$f.20513.0.0[dfsuicide$f.20513.0.0==3| dfsuicide$f.20513.0.0==4] <- 1
dfsuicide$f.20513.0.0[dfsuicide$f.20513.0.0==2] <- NA

dfsuicide$suicide_ind <- ifelse((dfsuicide$v20437==1 | dfsuicide$f.20513.0.0==1), 1, NA)
dfsuicide$suicide_ind[is.na(dfsuicide$suicide_ind) & dfsuicide$v20437==0 & dfsuicide$f.20513.0.0==0] <- 0

# md15 with suicide
dfsuicide$suicide <- ifelse((dfsuicide$suicide_ind ==1 & dfsuicide$md15==1), 2, NA)

dfsuicide$suicide[is.na(dfsuicide$suicide) & (dfsuicide$md15==1& dfsuicide$suicide_ind==0)] <- 1

dfsuicide$suicide[is.na(dfsuicide$suicide) & (dfsuicide$control_md==1)] <- 0

# cidi mdd with suicide
dfsuicide$suicide_cidi <- ifelse((dfsuicide$suicide_ind ==1 & dfsuicide$cidi_mdd==1), 2, NA)

dfsuicide$suicide_cidi[is.na(dfsuicide$suicide_cidi) & (dfsuicide$cidi_mdd==1& dfsuicide$suicide_ind==0)] <- 1

dfsuicide$suicide_cidi[is.na(dfsuicide$suicide_cidi) & (dfsuicide$control_md==1)] <- 0
```

### Select variables
```{r}
dfsuicide <- dfsuicide %>%
  select(c("fid", "eid", "suicide", "suicide_cidi"))
```

***************HHHHHH****************
##. 7 Impairment
Impact on normal roles during worst period of depression (ukb 20440)
20440
Mild (1):
20440 = 0 (not at all) / 1 (A little) 

moderate (2):
20440 = 2 (somewhat)

severe (3):
20440 = 3 (A lot)
```{r}
dfimpair <- rawcidi_ab %>%
  select(c("eid","f.20440.0.0")) %>%
  merge(., final_pop, by="eid")
names(dfimpair)[2] <- "impair"

# only case of MD15 were included in impair 123
dfimpair$impair[dfimpair$impair == 0] <- 1
dfimpair$impair[is.na(dfimpair$impair)& dfimpair$control_md] <- 0

# Note: Cannot derive impairment within cidi because 20440=3 is a criteria for cidi MDD
```

### Select variables
```{r}
dfimpair <- dfimpair %>%
  select(c("fid", "eid", "impair"))
```

***************HHHHHH****************
## 8. Postpartum
WOMEN who:
  1. Self reported post-natal depression UBK field 20002 code 1531
or  
  2. Diagnosed Mental and behavioural disorders associated with the puerperium, code F53 for UKB fields 41202 or 41202	
or
  3. Self-reported Depression possibly related to childbirth code 1 for UKB field 20445


### Compute variable based on criteria 1 20002 code 1531
Subset data criteria 1
```{r}
index20002 = grep(20002, names(raw180427), value = F)
df20002 <- select(raw180427,c(1,all_of(index20002)))
```

```{r}
# To list all the unique values in the data cols
disease=unique(as.vector(as.matrix(df20002[,2:88])))

# Check if any of the field 20002* has code 1531
df20002$code1531 <- as.numeric(apply(df20002[, 2:88], 1, function(x) any(x==1531)))

# Select columns
df20002_1531<- df20002 %>%
    select(c(eid, code1531))
```

# Subset data criteria 2 ICD = F53
```{r}
# Find the column index
index41202 = grep(41202, names(raw180427), value = F)
index41204 = grep(41204, names(raw180427), value = F)
# Subset data
df4120x <- select(raw180427,c(1,all_of(index41202),all_of(index41204)))
```

# Compute variable based on criteria 2,3 ICD = F53 and 20445 =1
```{r}
# To list all the unique values in the data cols
code_4120x = unique(as.vector(as.matrix(df4120x[,2:816])))

# Make a list of code F53 that exist in the data
f53_list = grep("F53", code_4120x, fixed=F, value = T)

# Check if there is any F53 code in each ID
df4120x$f53 <- as.numeric(apply(df4120x[, 2:816], 1, function(x) any(x %in% f53_list)))

df4120x$f53[df4120x$f53==0] <- NA

# Select column
dfppd<- df4120x %>%
  select(c(eid, f53)) %>%
  merge(., df20002_1531, by="eid") %>%
  merge(., raw20445, by="eid") %>%
  merge(., final_pop,by="eid") %>%
  merge(., rawsex, by="eid") %>%
  select(c(fid, eid, code1531, f53, f.20445.0.0, md15, cidi_mdd, control_ppd, sex))
```

### Combine 3 criteria
```{r}
dfppd$ppd_ind <- as.numeric(apply(dfppd[,3:5], 1, function(x) any(x==1)))

dfppd$ppd_ind[dfppd$ppd_ind==1 & dfppd$sex ==1] <- NA

# ppd within md15
dfppd$ppd<- ifelse((dfppd$md15==1 & dfppd$ppd_ind==1), 1, NA)
dfppd$ppd[is.na(dfppd$ppd) & dfppd$control_ppd ==1] <- 0

# ppd within cidi MDD
dfppd$ppd_cidi<- ifelse((dfppd$cidi_mdd==1 & dfppd$ppd_ind==1), 1, NA)
dfppd$ppd_cidi[is.na(dfppd$ppd_cidi) & dfppd$control_ppd ==1] <- 0
```
 
### Select variables

```{r}
dfppd <- dfppd %>%
  select(c(fid, eid, ppd, ppd_cidi))
```


